---
title: "Midterm_draft"
author: ""
date: "3/2/2021"
output: html_document
---

```{r}
drugs <- read.csv(file="salesweekly.csv", header=TRUE)

drugs$datum <- as.Date(drugs$datum, "%m/%d/%Y")
anhist <- drugs$R06
head(drugs)

#Citation: Chapter 05 notes, https://www.earthdatascience.org/courses/earth-analytics/time-series-data/date-class-in-r/

plot(anhist~drugs$datum, type='l')

```

The fluctuations in antihistamine sales appears to be cyclical. As antihistamines are frequently used to treat allergies, which can often be triggered by changes in the season [https://acaai.org/allergies/seasonal-allergies]. Let's take a look at this data in the frequency domain to see if our hypothesis of a yearly antihistamine sale cycle is supported:
```{r, echo=FALSE}
anhist_spec <- spectrum(anhist, plot=FALSE)
plot(anhist_spec$freq*52, anhist_spec$spec, type='l', log='y', main="Weekly Antihistamine Sale Spectrum", xlab="Cycles Per Year", ylab="spectrum")
```

There does appear to be a peak at approximately 1 cycle per year, but let's see if we can make the cyclical pattern even clearer by using a loess bandpass filter. We'll try to separate out the low frequency trend and high frequency noise to extract the cyclical behavior:

```{r}
a_low <- ts(loess(anhist~as.numeric(drugs$datum), span=0.5)$fitted, start=2014, frequency=52)
a_high <- ts(anhist - loess(anhist~as.numeric(drugs$datum), span=0.1)$fitted, start=2014, frequency=52)


a_cycles <- anhist - a_high - a_low

plot(ts.union(anhist, a_low, a_high, a_cycles))

##Citation: From Chapter 8 notes
```

The mid-range frequencies look like a fairly regular, one year cycle, which is promising. Let's plot a periodogram and extract the peak frequency of this cycle to see if that is borne out by the data:

```{r}
cycle_spec <- spectrum(a_cycles, main="Spectrum of Antihistamine Cycles", ylab="spectrum", xlab="Cycles per year")
abline(v=c(1, 2), col=c("red", "blue"), lty=c(2,2))

```

The above plot shows the periodogram of the mid-range frequency cyclic behavior, with the red line representing our expected peak frequency of 1 cycle per year, and the blue line representing the first harmonic of that expected frequency: 2 cycles per year.

Let's calculate the actual frequency of that first peak and the corresponding period:

```{r}
peak_freq <- cycle_spec$freq[which.max(cycle_spec$spec)]
peak_freq
period <- 1/peak_freq
period
```

The periodogram does indeed return a peak frequency of approximately 0.975, which corresponds to a period of approximately 1.03 years. This does indeed support the idea that antihistamine sales experience an annual cycle.
