---
title: "STATS 531 Midterm Project"
author: "Xingwen Wei"
date: "March 6, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Pharma Sales Dataset

Pharma sales dataset is built from transactional data collected in six years (2014-2019) indicating the date and time of the sale and pharmaceutical drug brand and quantity. The drugs are classified into eight groups by Anatomical Therapeutic Chemical (ATC) Classification System. For the purpose of this project, we focus on one specific group of drugs, R06, in which are antihistamines for systemic use. The main concern of this analysis is to find the connection between seasonal change and the difference in antihistamine sales across the year.

# 2 Exploratory Data Analysis
```{r, echo=F, warnings=F}
library(ggplot2)
library(tidyr)
library(GGally)
weekly <- read.csv("archive/salesweekly.csv", header=TRUE)
weekly$datum <- as.Date(weekly$datum, "%m/%d/%Y")
```

## 2.1 Basic Statistics

Firstly, we want to look at the basic statistics of the drug sales.
```{r, echo=F}
summary(weekly)
```

From the following boxplots, we find that all of the drug sales have outliers.
```{r, echo=F}
weekly_long <- gather(weekly, drug_type, quantity, M01AB:R06, factor_key = TRUE)
ggplot(weekly_long, aes(y=quantity)) + geom_boxplot(fill='slateblue', alpha=0.2) + facet_wrap(vars(drug_type), scales = 'free_y')
```

From the following pairwise scatterplot, we find that there is no sign of strong linear relationship between R06 and other variables
```{r, echo=F}
ggpairs(weekly[, 1:9])
```


According to the coefficient of time-related attribute in the linear model, the mean is barely increasing in time, although we cannot trust the model statistics because of the violation of assumptions to the error term based on the residual acf plot.
From the time plot of R06 sales against time, we find stationary model may not be appropriate for the data.
```{r, echo=F}
weekly$Day <- as.numeric(format(weekly$datum, format='%d'))
weekly$Month <- as.numeric(format(weekly$datum, format='%m'))
weekly$Year <- as.numeric(format(weekly$datum, format='%Y'))
weekly$time <- weekly$Year+(weekly$Month*31+weekly$Day)/372
lmod <- lm(R06~weekly$time+I(weekly$time^2), data=weekly)
summary(lmod)
acf(lmod$residuals)
Z <- cbind(1, weekly$time, weekly$time^2)
pred <- Z%*%lmod$coefficients
plot(R06~datum, data=weekly, type='l', xlab='date', main='weekly R06 sales')
lines(y=pred, x=weekly$datum, col='red')
```


## 2.2 Stationarility and Seasonality

From the sample autocorrelation plot, we find that the sample ACF has a peak at around 52, indicating strong evidence for period around 52, which corroberates with the inital assumption that the yearly period seasonality.
The periodic waves in sample ACF shows seasonal autoregressive model should be appropriate for this data.
```{r, echo=F}
acf(weekly$R06, lag=100, main='weekly R06 sample autocorrelation')
```

## 2.3 Frequency Domain Analysis

We can investigate the drug sale data from the frequency domain to get insights from another perspective.
We can examine the periodogram of the weekly sales data and decompose the R06 drug sale into trend, noise and cycle components.

The fluctuations in antihistamine sales appears to be cyclical. As antihistamines are frequently used to treat allergies, which can often be triggered by changes in the season ("Common Seasonal Allergy Triggers"). Let's take a look at this data in the frequency domain to see if our hypothesis of a yearly antihistamine sale cycle is supported:

```{r, echo=FALSE}
anhist <- weekly$R06
anhist_spec <- spectrum(anhist, plot=FALSE)
plot(anhist_spec$freq*52, anhist_spec$spec, type='l', log='y', main="Weekly Antihistamine Sale Spectrum", xlab="Cycles Per Year", ylab="spectrum")
```

There does appear to be a peak at approximately 1 cycle per year, but let's see if we can make the cyclical pattern even clearer by using a loess bandpass filter. We'll try to separate out the low frequency trend and high frequency noise to extract the cyclical behavior (Chapter 8 Notes):

```{r}
a_low <- ts(loess(anhist~as.numeric(weekly$datum), span=0.5)$fitted, start=2014, frequency=52)
a_high <- ts(anhist - loess(anhist~as.numeric(weekly$datum), span=0.1)$fitted, start=2014, frequency=52)


a_cycles <- anhist - a_high - a_low

plot(ts.union(anhist, a_low, a_high, a_cycles))

##Citation: From Chapter 8 notes
```

The mid-range frequencies look like a fairly regular, one year cycle, which is promising. Let's plot a periodogram and extract the peak frequency of this cycle to see if that is borne out by the data:

```{r}
cycle_spec <- spectrum(a_cycles, main="Spectrum of Antihistamine Cycles", ylab="spectrum", xlab="Cycles per year")
abline(v=c(1, 2), col=c("red", "blue"), lty=c(2,2))

```

The above plot shows the periodogram of the mid-range frequency cyclic behavior, with the red line representing our expected peak frequency of 1 cycle per year, and the blue line representing the first harmonic of that expected frequency: 2 cycles per year.

Let's calculate the actual frequency of that first peak and the corresponding period:

```{r}
peak_freq <- cycle_spec$freq[which.max(cycle_spec$spec)]
peak_freq
period <- 1/peak_freq
period
```

The periodogram does indeed return a peak frequency of approximately 0.975, which corresponds to a period of approximately 1.03 years. This does indeed support the idea that antihistamine sales experience an annual cycle.

# 3 Fitting an ARMA model

To establish a baseline model for comparison later, we start with an ARIMA model without considering the seasonal effects.

## 3.1 Model Selection for ARIMA Model

We choose to use an AIC table as our first step in the model selection process.
According to the AIC table below, ARMA(1, 1) has the best AIC value 2047.86.
We also spot an evidence for mathematical instability at ARMA(2, 2) and ARMA(3, 2) where the AIC value increased more than 2.

```{r, echo=F, warning=FALSE}
arma_aic_table <- function(data, P, Q){
  table <- matrix(NA, (P+1), (Q+1))
  for(p in 0:P){
    for(q in 0:Q){
      table[p+1, q+1] <- arima(data, order=c(p, 0, q))$aic
    }
  }
  dimnames(table) <- list(paste("AR", 0:P, sep=''), paste("MA", 0:Q, sep=''))
  table
}
r06_arma_aic_table <- arma_aic_table(weekly$R06, 3, 3)
require(knitr)
kable(r06_arma_aic_table, digits=2)
```

## 3.2 Model Diagnostics for ARIMA Model

We want to assess the goodness of fit of the ARMA(1, 1) model by looking at its residuals.
Based on the residual ACF, there is no strong evidence for autocorrelation between the residual values at different lags.
We superimposed the simulated data according to our ARMA(1, 1) model on top of the original R06 sales data.
As we expected, the ARMA model does not capture the seasonal effects and fits poorly.

```{r, echo=F}
arma11 <- arima(weekly$R06, order=c(1, 0, 1))
acf(arma11$residuals, main='ARMA(1, 1) ACF')
plot(R06~datum, data=weekly, type='l', xlab='date', main='weekly R06 sales')
t <- arima.sim(model=list(ar=0.902, ma=-0.3532), n=302, sd=sqrt(arma11$sigma2))+mean(weekly$R06)
lines(x=weekly$datum, y=t, col='red')
```

# 4 Fitting SARIMA Model

As we found from our exploratory data analysis, the time series data has a seasonal component with yearly period of 52 data points. Now we want to add the seasonal component into our model and see how much it improves the fit.

## 4.1 Model Selection for SARIMA Model



```{r, echo=F, warning=FALSE}
library(forecast)
sarma_aic_table <- function(data, P, Q){
  table <- matrix(NA, (P+1), (Q+1))
  for(p in 0:P){
    for(q in 0:Q){
      table[p+1, q+1] <- arima(data, order=c(p, 0, q), seasonal=list(order=c(1, 0, 0), period=52))$aic
    }
  }
  dimnames(table) <- list(paste("AR", 0:P, sep=''), paste("MA", 0:Q, sep=''))
  table
}

r06_sarma_aic_table <- sarma_aic_table(weekly$R06, 2, 2)
require(knitr)
kable(r06_sarma_aic_table, digits=2)

sarma11 <- arima(weekly$R06, order=c(1, 0, 1), seasonal=list(order=c(1, 0, 0), period=52))
t1 <- simulate(sarma11, nsim=302)
plot(R06~datum, data=weekly, type='l', xlab='date', main='weekly R06 sales')
lines(x=weekly$datum, y=t1, col='red')
```



## 4.2 Model Diagnostics for SARIMA Model

# 5 Conclusion



# References and Group Activity

We worked collaboratively as a group. There was a lot of group consultation and some overlap in individual investigations and modelling of data. Here, broadly is what we each worked on individually:

Jessica Leviton: Frequency domain analysis, including decomposition into different signal components and extraction of the dominant frequency of the antihistamine sale cycle.

The following sources to compose this project:

“Seasonal Allergies.” ACAAI Public Website, 29 Oct. 2018, acaai.org/allergies/seasonal-allergies.

Ionides, Edward L. "Chapter 8: Smoothing in the Time and Frequency Domains." 2021, ionides.github.io/531w21/08/notes.pdf. 
